---
title: "Linear Regression Model Project"
author: "Jason Barnes"
date: "5/21/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Linear Regression Model 
###(Bikeshare Challenge)

You are provided hourly rental data spanning two years. For this competition, the training set is comprised of the first 19 days of each month, while the test set is the 20th to the end of the month. You must predict the total count of bikes rented during each hour covered by the test set, using only information available prior to the rental period.

####Libraries 
```{r}
library(tidyverse)
library(ggthemes)
library(lubridate)
library(knitr)
library(kableExtra)
#devtools::install_github('yihui/tinytex')
library(tinytex)
library(rmarkdown)
```

####Data
Data can be downloaded at https://www.kaggle.com/c/bike-sharing-demand/data#
```{r}
bike <-  read.csv("bikeshare.csv")
```

####Explore the data
```{r}
headbike <- head(bike)
kable(headbike) %>% 
    kable_styling(bootstrap_options = "striped", full_width = FALSE)
```
#####Check the structure of bike
```{r}
str(bike)
```
#####Factors
Since weather and season are factors we will want to change the data to as.factor

```{r}
bike$season <- as.factor(bike$season)
bike$weather <- as.factor(bike$weather)
```
Now we will call structure aagin so we can verify the changes
```{r}
str(bike)
```
#####Visualize the data
count vs temp
```{r}
ggplot(bike, aes(temp, count)) +
    geom_point(aes(color = temp), alpha = .5) +
    labs(x = "Temperature", y = "Number of bike rentals", title = "Rentals v Temp") +
    theme_clean()

```

In order to vizualize Need to turn datetime into POSIXct data field
```{r}
bike$datetime <- as.POSIXct(bike$datetime)
```

count vs datetime
```{r}
ggplot(bike, aes(datetime, count)) +
    geom_point(aes(color = temp)) +
    scale_color_continuous(low = "turquoise2", high = "orange") +
    labs(x = "Date", y = "Number of bike rentals", title = "Rentals vs Date") +
    theme_clean()
```

Coorelation between number of bikes rented and the temperature 
```{r}
cor(bike[,c("temp","count")])
```

Expolore seasonality with a boxplot
```{r}
ggplot(bike, aes(season, count)) +
    geom_boxplot(aes(color = season)) +
    theme_clean()
```

Add an hour column 
```{r}
bike$hour <- sapply(bike$datetime, function(x){format(x,"%H")})
```

Segment data into weekday and weekend
```{r}
workday <- bike %>% 
    filter(workingday == 1)
nonworkday <- bike %>% 
    filter(workingday == 0)
```

Visulaize weekday
```{r}
ggplot(workday, aes(hour, count)) +
    geom_point(aes(color = temp), position = position_jitter(width = 1, height = 0)) +
    scale_color_gradientn(colours = c('dark blue','blue','light blue','light green','yellow','orange','red')) +
    labs(title = "Weekday; Rentals by Hour") +
    theme_clean()
```

>Notice the peak in the morning before work and the evening after work?

Visualize the weekend 
```{r}
ggplot(nonworkday, aes(hour, count)) +
    geom_point(aes(color = temp), position = position_jitter(width = 1, height = 0)) +
    scale_color_gradient2_tableau(palette = "Red-Green Diverging") +
    labs(title = "Weekend; Rentals by Hour") +
    theme_clean()
```

#####Building the Model with just temperature as an attribute
```{r}
temp.model <- lm(formula = count ~ temp, data = bike )
summary(temp.model)
```
#####Interpeting the Model
>Interpreting the intercept: It is the value of y when x=0.Thus, it is the estimated number of rentals when the temperature is 0 degrees Celsius. Note: It does not always make sense to interpret the intercept. 


>Interpreting the "temp" coefficient: It is the change in y divided by change in x, or the "slope". Thus, a temperature increase of 1 degree Celsius is associated with a rental increase of 9.17 bikes. This is not a statement of causation. Coefficient would be negative if an increase in temperature was associated with a decrease in rentals.

#####Predict the number of rentals if the temperature was 25 degrees Celsius
```{r}
# y = mx + b
9.17 * (25) + 6.0462
```

##### Building the final model 
* Attributes
    + season
    + holiday
    + workingday
    + weather
    + temp
    + humidity
    + windspeed
    + hour (factor)

Convert hour to a numeric value
```{r}
bike$hour <- sapply(bike$hour, as.numeric)
```

Model 
```{r}
model_bike <- lm(count ~ . -casual - registered - datetime - atemp, bike)
summary(model_bike)
```





