---
title: "Linear Regression"
date: "5/18/2019"
---
###Pull in the data set
```{r df}
df <- read.csv("student-mat.csv", sep = ';')
```
###Get to know your data
```{r}
head(df)
```
```{r}
summary(df)
```
###Is there any null values
```{r}
any(is.na(df))
```
###Structure 
Make sure the factors with appropriate levels where imported.
```{r}
str(df)
```

Identify what columns have numeric values

```{r}
num.cols <- sapply(df, is.numeric)
```
###Coorelation
Build a correlation table 
```{r}
cor.data <- cor(df[,num.cols])
```
View the correlation table
```{r}
cor.data
```
As you can see it is pretty difficult to see what is happening, so lets visualize it will coorplot.
```{r}
library(corrplot)
corrplot(cor.data, method = "color")
```

You can also use corrgram for a different perspective.

```{r}
library(corrgram)
corrgram(df)
```
```{r}
corrgram(df, order = TRUE, lower.panel = panel.shade, upper.panel = panel.pie, text.panel = panel.txt)
```
```{r}
library(ggplot2)
```
Lets look at a histogram of the 3rd quarter test scores (G3) after all this is what we are trying to perdict
```{r}
ggplot(df, aes(G3)) +
    geom_histogram(bins = 20, alpha = .5, fill = "blue")
```


